---
title: "Alternative"
---

Heat map of gene expression in PROGNOSTIC collection ER-/HER2-

```{r}
set.seed(123)

load(file = '../data/Chemo.rda')

data <- chemo_normalized
expression <- as.data.frame(t(exprs(data)))

expression$study <- chemo_normalized$StudyID
expression <- rbind(expression[expression$study == "GSE41998",],expression[expression$study == "GSE16446",], expression[expression$study == "E_MTAB_365",])

dim(expression)

CTM <- c("CXCL13","PRF1", "IRF1", "IKZF1", "GZMB", "HLA-E")

ER_metagene <- c("CCDC170", "ESR1", "ABAT", "SLC39A6", "GATA3", "SCUBE2")
HER_metagene <- c("ERBB2", "PGAP3", "STARD3", "GRB7", "PNMT", "PSMD3", "GSDMB", "RPL19", "FGFR4", "CAP1")
threshold_ER <- 7.5
threshold_HER <- 8.35

mean_ER <- rowMeans(expression[, ER_metagene])
mean_HER <- rowMeans(expression[, HER_metagene])

ER_status <- mean_ER
ER_status[mean_ER > threshold_ER] <- T; ER_status[mean_ER <= threshold_ER] <- F
HER_status <- mean_HER
HER_status[mean_HER > threshold_HER] <- T; HER_status[mean_HER <= threshold_HER] <- F

samples_doubleneg <- names(ER_status[(ER_status + HER_status) == 0])
exp_doubleneg <- data.frame(expression[samples_doubleneg, CTM])
samples_HERpos <- names(HER_status[HER_status == 1])
exp_HERpos <- expression[samples_HERpos, CTM]
samples_HERneg <- names(HER_status[HER_status == 0])
exp_HERneg <- expression[samples_HERneg, CTM]

dim(exp_doubleneg)
```

```{r}
metagene <- c("CXCL13","PRF1", "IRF1", "IKZF1", "GZMB", "HLA.E")
#Find tertiles, assign low,medium,high to each group
sum <- rowMeans(exp_doubleneg[,metagene])
tert = quantile(sum, c(0:3/3))
exp_doubleneg$group = with(exp_doubleneg, cut(sum, tert, include.lowest = T, labels = c("Low", "Medium", "High")))
```

Generate heat map of expression of CTM
```{r}
library(ComplexHeatmap)
Heatmap(exp_doubleneg[,metagene], cluster_rows = F, name = "ER-/HER-",cluster_columns = F, split = exp_doubleneg$group, show_row_names = F)
```


Alternative clustering method

Load data.
```{r}
library("Biobase")
library(WGCNA)
options(stringsAsFactors = FALSE)

set.seed(123)
load(file = '../data/Generic.rda')

data <- generic_normalized
expression <- t(exprs(data))

```

Choose the soft-thresholding powers
```{r}
# Choose a set of soft-thresholding powers
powers = c(c(1:10), seq(from = 12, to=20, by=2))
# Call the network topology analysis function
sft = pickSoftThreshold(expression, powerVector = powers, verbose = 5)
# Plot the results:
sizeGrWindow(9, 5)
par(mfrow = c(1,2));
cex1 = 0.9;
# Scale-free topology fit index as a function of the soft-thresholding power
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
main = paste("Scale independence"));
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
labels=powers,cex=cex1,col="red");
# this line corresponds to using an R^2 cut-off of h
abline(h=0.90,col="red")
# Mean connectivity as a function of the soft-thresholding power
plot(sft$fitIndices[,1], sft$fitIndices[,5],
xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
```
We choose power 4, which is the lowest power to reach to curve at about 0.9.
The min cluster size is 20 as in the original analysis.

Define module and identify modules. Use default of function.
```{r}
net = blockwiseModules(expression, power = 4,
TOMType = "unsigned", minModuleSize = 20,
reassignThreshold = 0, mergeCutHeight = 0.25,
numericLabels = TRUE, pamRespectsDendro = FALSE,
verbose = 3)
```

Let's see the identified modules of co-expressed genes.
```{r}
#Print number of clusters and their size
table(net$colors)
moduleColors = labels2colors(net$colors)
colors <- unique(moduleColors)

clusters <- vector(mode='list', length=length(colors))
for (i in 1:length(colors))
{
modGenes = (moduleColors==colors[i])
clusters[[i]] <- colnames(expression[,modGenes])
}
```
```{r}
clusters[[16]]
```

GO enrichment analysis
```{r}
library(enrichR)
setEnrichrSite("Enrichr") # Human genes
websiteLive <- TRUE

dbs <- c("GO_Molecular_Function_2015", "GO_Cellular_Component_2015", "GO_Biological_Process_2015")
if (websiteLive) {
    enriched <- enrichr(clusters[[16]], dbs)
}
if (websiteLive) plotEnrich(enriched[[3]], showTerms = 20, numChar = 40, y = "Count", orderBy = "P.value")

```

Refinement of metagene in the ER-HER- subtype
```{r}
load(file = '../data/Prognostic.rda')

immune_metagene <- clusters[[16]]

data <- prognostic_normalized
expression <- as.data.frame(t(exprs(data)))

outcome <- prognostic_normalized$t.dmfs.norm.years
expression$DMFSVal <- outcome
expression$DMFSVal[expression$DMFSVal > 5] <- 5 
outcome[outcome <= 5] <- T; outcome[outcome > 5] <- F
expression$DMFS <- outcome

dim(expression)

#Define ER/HER2 status
ER_metagene <- c("CCDC170", "ESR1", "ABAT", "SLC39A6", "GATA3", "SCUBE2")
HER_metagene <- c("ERBB2", "PGAP3", "STARD3", "GRB7", "PNMT", "PSMD3", "GSDMB", "RPL19", "FGFR4", "CAP1")
threshold_ER <- 7.5
threshold_HER <- 8.35

mean_ER <- rowMeans(expression[, ER_metagene])
mean_HER <- rowMeans(expression[, HER_metagene])

ER_status <- mean_ER
ER_status[mean_ER > threshold_ER] <- T; ER_status[mean_ER <= threshold_ER] <- F

HER_status <- mean_HER
HER_status[mean_HER > threshold_HER] <- T; HER_status[mean_HER <= threshold_HER] <- F

samples_doubleneg <- names(ER_status[(ER_status + HER_status) == 0])
exp_doubleneg <- data.frame(expression[samples_doubleneg, clusters[[16]]])
exp_doubleneg$DMFSVal <- expression[samples_doubleneg,]$DMFSVal
exp_doubleneg$DMFS <- expression[samples_doubleneg,]$DMFS 

dim(exp_doubleneg)
```

```{r}
source('refinement_func.R')#Load func for the refinment process

metagene <- clusters[[16]]; metagene[metagene == "HLA-E"] <- "HLA.E"; metagene[metagene== "HLA-DMB"] <- "HLA.DMB"; metagene[metagene== "HLA-DOB"] <- "HLA.DOB"; metagene[metagene== "HLA-DPA1"] <- "HLA.DPA1"; metagene[metagene== "HLA-DPB1"] <- "HLA.DPB1";metagene[metagene== "HLA-DQB1"] <- "HLA.DQB1";metagene[metagene== "HLA-DRA"] <- "HLA.DRA";metagene[metagene== "HLA-DRB1"] <- "HLA.DRB1" ;metagene[metagene== "HLA-G"] <- "HLA.G"; metagene[metagene== "HLA-DMA"] <- "HLA.DMA";

#Perform analysis 100 and take mean of DMFS percentage and p-values
mean_score <- c(0)
for (i in 1:10){
  mean_score <- mean_score + refine_metagene(exp_doubleneg, metagene)
}

mean_score <- mean_score/10
```

```{r}
#Get best number of genes -> best mean p-value
N <- 3*(length(metagene) - 9)
DMFS_group <- mean_score[1:N]; p_value <- mean_score[N: (N+ N/3)]

min_index = which.min(p_value)
print(paste("The best N is: ", min_index + 9, ", with p-value of ", min(p_value)))
DMFS_group[((min_index*3) - 2):(min_index*3)]

#ranked_genes <- univariate_Cox(exp_doubleneg, metagene)
#best_genes <- names(ranked_genes)[0:(min_index + 9)]
```

Validation in the CHEMO collection by Keipler meier analysis
```{r}
load(file = '../data/Chemo.rda')

data <- chemo_normalized
expression <- as.data.frame(t(exprs(data)))

outcome <- chemo_normalized$t.dmfs.norm.years
expression$DMFSVal <- outcome
expression$DMFSVal[expression$DMFSVal > 5] <- 5 
outcome[outcome <= 5] <- T; outcome[outcome > 5] <- F
expression$DMFS <- outcome


expression$study <- chemo_normalized$StudyID

#For survival figure 2B, only Patel + desmel dataset
expression <- rbind(expression[expression$study == "E_MTAB_365",], expression[expression$study == "GSE16446",])

dim(expression)

#Define ER/HER2 status of samples 
CTM <- c("SOCS1","NKG7","CXCR6","CTSW","GZMH","IL2RB","PTPN7","C2","APOE","NM_006847","IKZF1","PLEK","GM2A","CYBA","IL2RG","RAC2","NR1H3","CCND2","IFNG","PRF1","IL21R","PTPN22","APOL3","IRF1","SIRPG","SLAMF7","ITGAL","RNASET2","HLA-DMA","MAN2B1","LAG3","SELPLG","C1QA","CECR1","CCR5","MZT2B", "DMFSVal", "DMFS")

ER_metagene <- c("CCDC170", "ESR1", "ABAT", "SLC39A6", "GATA3", "SCUBE2")
HER_metagene <- c("ERBB2", "PGAP3", "STARD3", "GRB7", "PNMT", "PSMD3", "GSDMB", "RPL19", "FGFR4", "CAP1")
threshold_ER <- 7.5
threshold_HER <- 8.35

mean_ER <- rowMeans(expression[, ER_metagene])
mean_HER <- rowMeans(expression[, HER_metagene])

ER_status <- mean_ER
ER_status[mean_ER > threshold_ER] <- T; ER_status[mean_ER <= threshold_ER] <- F
HER_status <- mean_HER
HER_status[mean_HER > threshold_HER] <- T; HER_status[mean_HER <= threshold_HER] <- F

samples_doubleneg <- names(ER_status[(ER_status + HER_status) == 0])
exp_doubleneg <- data.frame(expression[samples_doubleneg,CTM])

dim(exp_doubleneg)
```

Assign low, medium, high expression tertiles
```{r}
CTM <- best_genes #Change metagene here

exp_doubleneg$sum <- rowMeans(exp_doubleneg[,CTM])
tert = quantile(exp_doubleneg$sum, c(0:3/3))
exp_doubleneg$group = with(exp_doubleneg, cut(sum, tert, include.lowest = T, labels = c("Low", "Medium", "High")))
```

Kepler Meir analysis
```{r}
all_data <- exp_doubleneg

#Find tertiles, assign low,medium,high to each group
all_data$sum <- rowMeans(all_data[,CTM])
tert = quantile(all_data$sum, c(0:3/3))
all_data$group = with(all_data, cut(sum, tert, include.lowest = T, labels = c("Low", "Medium", "High")))

DMFS <- data.frame(all_data$DMFS, row.names = rownames(all_data)); colnames(DMFS) <- "event"
DMFS.Val <- data.frame(all_data$DMFSVal, row.names = rownames(all_data)); colnames(DMFS.Val) <- 'time'

#log-rank test
DMFS <- data.frame(all_data$DMFS, row.names = rownames(all_data))
DMFS.Val <- data.frame(all_data$DMFSVal, row.names = rownames(all_data))
fit <- survfit(Surv(DMFSVal, DMFS)~ group, data = all_data)
print(autoplot(fit,  conf.int = F, ylim = c(0, 1), xlab = 'Time(years)', ylab = "DMFS(proportion)"))
surv_diff <- survdiff(Surv(DMFSVal, DMFS)~ group, data = all_data)
p_value <- 1 - pchisq(surv_diff$chisq, length(surv_diff$n)-1, lower.tail = TRUE)

print(p_value)
```
