---
title: "Immune prognosis"
---

Load data all relevant data. Comment the appropriate selection of studies depending on the analysis performed.

```{r}

set.seed(123)
library("survival")
library(survminer)
library(ggfortify)

load(file = '../data/Chemo.rda')

data <- chemo_normalized
expression <- as.data.frame(t(exprs(data)))

outcome <- chemo_normalized$t.dmfs.norm.years
expression$DMFSVal <- outcome
expression$DMFSVal[expression$DMFSVal > 5] <- 5 
outcome[outcome <= 5] <- T; outcome[outcome > 5] <- F
expression$DMFS <- outcome

expression$pcr <- chemo_normalized$pCR

#For table 3, only ... datasets
#As binary Node pos vs neg
node <- chemo_normalized$node
N0 <- levels(node)[3]
node[node != N0] <- 1 ; node[node == N0] <- 0;
expression$node <- node
expression$node <- droplevels(expression$node)

#As binary grade III vs grade I and II
grade <- chemo_normalized$grade
grade[grade == 1] <- F; grade[grade == 2]  <- F; grade[grade == 3] <- T;
expression$grade <- grade

#As binary age > 50 vs age <= 50
age <- chemo_normalized$age
age[age <= 50] <- F; age[age > 50] <- T;
expression$age <- age

expression$study <- chemo_normalized$StudyID
#For logistic regression analysis
#expression <- rbind(expression[expression$study == "GSE41998",], expression[expression$study == "GSE16446",])

#For survival figure 2B, only Patel + desmel dataset
#expression <- rbind(expression[expression$study == "E_MTAB_365",], expression[expression$study == "GSE16446",])

#For multivariate COX analysis 
expression <- rbind(expression[expression$study == "GSE41998",],expression[expression$study == "GSE16446",], expression[expression$study == "E_MTAB_365",])

dim(expression)
```

Find ER/HER status. Define ER-HER- and HER2+ expression set. Load specific studies.

```{r}
CTM <- c("CXCL13","PRF1", "IRF1", "IKZF1", "GZMB", "HLA-E", "DMFSVal", "DMFS","pcr","node", "grade", "age")

ER_metagene <- c("CCDC170", "ESR1", "ABAT", "SLC39A6", "GATA3", "SCUBE2")
HER_metagene <- c("ERBB2", "PGAP3", "STARD3", "GRB7", "PNMT", "PSMD3", "GSDMB", "RPL19", "FGFR4", "CAP1")
threshold_ER <- 7.5
threshold_HER <- 8.35

mean_ER <- rowMeans(expression[, ER_metagene])
mean_HER <- rowMeans(expression[, HER_metagene])

ER_status <- mean_ER
ER_status[mean_ER > threshold_ER] <- T; ER_status[mean_ER <= threshold_ER] <- F
HER_status <- mean_HER
HER_status[mean_HER > threshold_HER] <- T; HER_status[mean_HER <= threshold_HER] <- F

samples_doubleneg <- names(ER_status[(ER_status + HER_status) == 0])
exp_doubleneg <- data.frame(expression[samples_doubleneg, CTM])
samples_HERpos <- names(HER_status[HER_status == 1])
exp_HERpos <- expression[samples_HERpos, CTM]
samples_HERneg <- names(HER_status[HER_status == 0])
exp_HERneg <- expression[samples_HERneg, CTM]

dim(exp_doubleneg); dim(exp_HERpos); dim(exp_HERneg)
```

Assign low, medium, high expression of CTM

```{r}
CTM <- c("CXCL13","PRF1", "IRF1", "IKZF1", "GZMB", "HLA.E")

exp_doubleneg$sum <- rowMeans(exp_doubleneg[,CTM])
tert = quantile(exp_doubleneg$sum, c(0:3/3))
exp_doubleneg$group = with(exp_doubleneg, cut(sum, tert, include.lowest = T, labels = c("Low", "Medium", "High")))

names(exp_HERpos)[names(exp_HERpos) == "HLA-E"] <- "HLA.E"
exp_HERpos$sum <- rowMeans(exp_HERpos[,CTM])
tert = quantile(exp_HERpos$sum, c(0:3/3))
exp_HERpos$group = with(exp_HERpos, cut(sum, tert, include.lowest = T, labels = c("Low", "Medium", "High")))

```

Logistic regression, compute odd ratio, ect.. Odd ratio = exp(coeff), CI = exp(coeff+/-1.96\*SE(coeff))

```{r}
logis_doubleneg <-glm(pcr ~ group, family=binomial, data=exp_doubleneg)
summary(logis_doubleneg)
```

```{r}
logis_HERpos <-glm(pcr ~ group, family=binomial, data=exp_HERpos)
summary(logis_HERpos)
confint(logis_HERpos)

```

```{r}
# Cochrane data from the 'rmeta'-package
base_data <- tibble(mean  = c(3.56, 3.03, 7.50, 8.99), 
                    lower = c(1.69, 1.29, 0.79, 0.93),
                    upper = c(8.39, 7.09, 71.03, 86.56),
                    study = c("Auckland", "Block", "Doran", "ff"),
                    deaths_steroid = c("36", "1", "4", "14"),
                    deaths_placebo = c("60", "5", "11", "20"),
                    OR = c("0.58", "0.16", "0.25", "0.70"))

summary <- tibble(mean  = 0.531, 
                  lower = 0.386,
                  upper = 0.731,
                  study = "Summary",
                  OR = "0.53",
                  summary = TRUE)

header <- tibble(study = c("", "Study"),
                 summary = TRUE)

empty_row <- tibble(mean = NA_real_)

cochrane_output_df <- bind_rows(header,
                                base_data,
                                empty_row)

cochrane_output_df %>% 
  forestplot(labeltext = c(study, deaths_steroid, deaths_placebo, OR), 
             xlog = TRUE, 
             col = fpColors(box = "royalblue",
                            line = "darkblue",
                            summary = "royalblue"))
```


Define and test the consensus T-cell related metagene for double negative and HER+

```{r}
CTM <- c("CXCL13","PRF1", "IRF1", "IKZF1", "GZMB", "HLA.E")
all_data <- rbind(exp_doubleneg,exp_HERpos)

#Find tertiles, assign low,medium,high to each group
all_data$sum <- rowMeans(all_data[,CTM])
tert = quantile(all_data$sum, c(0:3/3))
all_data$group = with(all_data, cut(sum, tert, include.lowest = T, labels = c("Low", "Medium", "High")))

DMFS <- data.frame(all_data$DMFS, row.names = rownames(all_data)); colnames(DMFS) <- "event"
DMFS.Val <- data.frame(all_data$DMFSVal, row.names = rownames(all_data)); colnames(DMFS.Val) <- 'time'


#log-rank test
DMFS <- data.frame(all_data$DMFS, row.names = rownames(all_data))
DMFS.Val <- data.frame(all_data$DMFSVal, row.names = rownames(all_data))
fit <- survfit(Surv(DMFSVal, DMFS)~ group, data = all_data)
print(autoplot(fit,  conf.int = F, ylim = c(0, 1), xlab = 'Time(years)', ylab = "DMFS(proportion)"))
surv_diff <- survdiff(Surv(DMFSVal, DMFS)~ group, data = all_data)
p_value <- 1 - pchisq(surv_diff$chisq, length(surv_diff$n)-1, lower.tail = TRUE)

print(p_value)
```

Only double_neg

```{r}
CTM <- c("CXCL13","PRF1", "IRF1", "IKZF1", "GZMB", "HLA.E")
all_data <- exp_doubleneg

#Find tertiles, assign low,medium,high to each group
all_data$sum <- rowMeans(all_data[,CTM])
tert = quantile(all_data$sum, c(0:3/3))
all_data$group = with(all_data, cut(sum, tert, include.lowest = T, labels = c("Low", "Medium", "High")))

DMFS <- data.frame(all_data$DMFS, row.names = rownames(all_data)); colnames(DMFS) <- "event"
DMFS.Val <- data.frame(all_data$DMFSVal, row.names = rownames(all_data)); colnames(DMFS.Val) <- 'time'

#log-rank test
DMFS <- data.frame(all_data$DMFS, row.names = rownames(all_data))
DMFS.Val <- data.frame(all_data$DMFSVal, row.names = rownames(all_data))
fit <- survfit(Surv(DMFSVal, DMFS)~ group, data = all_data)
print(autoplot(fit,  conf.int = F, ylim = c(0, 1), xlab = 'Time(years)', ylab = "DMFS(proportion)"))
surv_diff <- survdiff(Surv(DMFSVal, DMFS)~ group, data = all_data)
p_value <- 1 - pchisq(surv_diff$chisq, length(surv_diff$n)-1, lower.tail = TRUE)

print(p_value)
```

Only HER+

```{r}
CTM <- c("CXCL13","PRF1", "IRF1", "IKZF1", "GZMB", "HLA.E")
names(exp_HERpos)[names(exp_HERpos) == "HLA-E"] <- "HLA.E"
all_data <- exp_HERpos

#Find tertiles, assign low,medium,high to each group
all_data$sum <- rowMeans(all_data[,CTM])
tert = quantile(all_data$sum, c(0:3/3))
all_data$group = with(all_data, cut(sum, tert, include.lowest = T, labels = c("Low", "Medium", "High")))

DMFS <- data.frame(all_data$DMFS, row.names = rownames(all_data)); colnames(DMFS) <- "event"
DMFS.Val <- data.frame(all_data$DMFSVal, row.names = rownames(all_data)); colnames(DMFS.Val) <- 'time'

#log-rank test
DMFS <- data.frame(all_data$DMFS, row.names = rownames(all_data))
DMFS.Val <- data.frame(all_data$DMFSVal, row.names = rownames(all_data))
fit <- survfit(Surv(DMFSVal, DMFS)~ group, data = all_data)
print(autoplot(fit,  conf.int = F, ylim = c(0, 1), xlab = 'Time(years)', ylab = "DMFS(proportion)"))
surv_diff <- survdiff(Surv(DMFSVal, DMFS)~ group, data = all_data)
p_value <- 1 - pchisq(surv_diff$chisq, length(surv_diff$n)-1, lower.tail = TRUE)

print(p_value)
```

Multivariate cox regression for ER-/HER- patients

```{r}
CTM <- c("CXCL13","PRF1", "IRF1", "IKZF1", "GZMB", "HLA.E")
all_data <- exp_doubleneg

#Find tertiles, assign low,medium,high to each group
all_data$sum <- rowMeans(all_data[,CTM])
tert = quantile(all_data$sum, c(0:3/3))
all_data$group = with(all_data, cut(sum, tert, include.lowest = T, labels = c("Low", "Medium", "High")))

res.cox <- coxph(Surv(DMFSVal, DMFS) ~ group + node + grade, data =  all_data)
summary(res.cox)

```
