---
title: "Metagene refinement"
output: html_document
---
Define ER and HER metagene for status determination and immune metagene to be refined. Load data.
```{r}
set.seed(123)

load(file = '../data/Prognostic.rda')
ER_metagene <- c("CCDC170", "ESR1", "ABAT", "SLC39A6", "GATA3", "SCUBE2")
HER_metagene <- c("ERBB2", "PGAP3", "STARD3", "GRB7", "PNMT", "PSMD3", "GSDMB", "RPL19", "FGFR4", "CAP1")
threshold_ER <- 7.5
threshold_HER <- 8.35

immune_metagene <- c("CXCL13","PLEK","IFNG","SLAMF7","IL2RB","PRF1","IRF1","PTPN22","IKZF1","APOBEC3G","IL2RA","ITGAL","CXCL9","GZMA","GZMB","HLA-E","CCR5","CD8A","SIRPG","CST7","GNLY","CECR1","PNOC","LCP1","HLA-DMB","GIMAP5","SEL1L3", "FGL2", "BIN2")

print(length(immune_metagene))

data <- prognostic_normalized
expression <- as.data.frame(t(exprs(data)))

outcome <- prognostic_normalized$t.dmfs.norm.years
expression$DMFSVal <- outcome
expression$DMFSVal[expression$DMFSVal > 5] <- 5 
outcome[outcome <= 5] <- T; outcome[outcome > 5] <- F
expression$DMFS <- outcome

dim(expression)

```

Find ER/HER status. Define ER-HER- and HER2+ expression set.
```{r}
mean_ER <- rowMeans(expression[, ER_metagene])
mean_HER <- rowMeans(expression[, HER_metagene])

ER_status <- mean_ER
ER_status[mean_ER > threshold_ER] <- T; ER_status[mean_ER <= threshold_ER] <- F

HER_status <- mean_HER
HER_status[mean_HER > threshold_HER] <- T; HER_status[mean_HER <= threshold_HER] <- F

samples_doubleneg <- names(ER_status[(ER_status + HER_status) == 0])
exp_doubleneg <- data.frame(expression[samples_doubleneg, immune_metagene])
exp_doubleneg$DMFSVal <- expression[samples_doubleneg,]$DMFSVal
exp_doubleneg$DMFS <- expression[samples_doubleneg,]$DMFS 

samples_HERpos <- names(HER_status[HER_status == 1])
exp_HERpos <- expression[samples_HERpos, immune_metagene]
exp_HERpos$DMFSVal <- expression[samples_HERpos,]$DMFSVal 
exp_HERpos$DMFS <- expression[samples_HERpos,]$DMFS 

dim(exp_doubleneg); dim(exp_HERpos)

```

Cross-validated number of genes in metagene
```{r}
library("survival")
library(survminer)
library(ggfortify)

univariate_Cox <- function(test_data, genes){
  ranked_genes <- vector(mode="double", length=length(genes)); names(ranked_genes) <- genes
  for (i in 1:length(genes)){
    formula <- as.formula(paste('Surv(DMFSVal, DMFS, type="right")~', genes[i]))
    res.cox <- coxph(formula, data = test_data)
    x <- summary(res.cox)
    ranked_genes[i] <- x$wald["pvalue"]
  }
  ranked_genes <- sort(ranked_genes, decreasing = F)
  
  return (ranked_genes)
}

```

```{r}
compute_DMFS <- function(exp){
  #Compute percentage DMFS for each group
  DMFS.low <- (sum(exp[exp$group == 1,]$DMFS)/length(exp[exp$group == 1,]$DMFS))
  DMFS.medium <- (sum(exp[exp$group == 2,]$DMFS)/length(exp[exp$group == 2,]$DMFS))
  DMFS.high <- (sum(exp[exp$group == 3,]$DMFS)/length(exp[exp$group == 3,]$DMFS))
  
  DMFS.groups <- c(DMFS.low,DMFS.medium,DMFS.high)
  DMFS.groups <- 1- DMFS.groups 
  DMFS.groups[is.na(DMFS.groups)] <- 0
  
  return (DMFS.groups)
}

assign_exp_group<- function(exp){
  #Assign low,medium,high to each sample
  exp$sum <- rowMeans(exp)
  tert = quantile(exp$sum, c(0:3/3))
  exp$group = with(exp, cut(sum, tert, include.lowest = T, labels = c("Low", "Medium", "High")))

return (exp$group)
}
```

```{r}
refine_metagene<- function(exp, genes){
  #Shuffle the data and create fold
  exp <- exp[sample(nrow(exp)),]
  folds <- cut(seq(1,nrow(exp)),breaks=10,labels=FALSE)
  
  #Perform 10 fold CV
  exp_group <- data.frame(matrix(0, nrow = nrow(exp), ncol = length(genes)-9))
  colnames(exp_group) <- 10:length(genes); rownames(exp_group) <- rownames(exp)
  for(i in 1:10){
    #Get train and test data sets
    testIndexes <- which(folds==i,arr.ind=TRUE)
    testData <- exp[testIndexes, ]
    trainData <- exp[-testIndexes, ]
      
    #Perform univariate Cox analysis of each gene on train set
    ranked_genes <- univariate_Cox(trainData, genes)

    #Assign group to test sets for each metagene size
    for(N in 10:length(genes)){
      metagene <- names(ranked_genes)[1:N]
      exp_group[testIndexes, N-9] <- assign_exp_group(testData[, metagene])
    }
  }
  
  #Compute 5 years DMFS (%) and log-rank test for each size of metagene
  DMFS_metagenes = c(); p_value = c()
  for(N in 10:length(genes)){
    data <- data.frame(exp_group[N-9], exp$DMFS, exp$DMFSVal, row.names = rownames(exp))
    colnames(data) <- c("group","DMFS", "time")
    
    #Compute 5 years DMFS (%)
    DMFS_metagenes <- append(DMFS_metagenes, compute_DMFS(data))
  
    #Log-rank test
    surv_diff <- survdiff(Surv(time, DMFS)~ group, data = data)
    p_value <- append(p_value, (1 - pchisq(surv_diff$chisq, length(surv_diff$n)-1, lower.tail = TRUE)))
  }
  
  return(c(DMFS_metagenes, p_value))
}
```

Refine metagene for double negative patients
```{r}
immune_metagenes <- c("CXCL13","PLEK","IFNG","SLAMF7","IL2RB","PRF1","IRF1","PTPN22","IKZF1","APOBEC3G","IL2RA","ITGAL","CXCL9","GZMA","GZMB","HLA.E","CCR5","CD8A","SIRPG","CST7","GNLY","CECR1","PNOC","LCP1","HLA.DMB","GIMAP5","SEL1L3", "FGL2", "BIN2")

#Perform analysis 100 and take mean of DMFS percentage and p-values
mean_score <- c(0)
for (i in 1:100){
  mean_score <- mean_score + refine_metagene(exp_doubleneg, immune_metagenes)
}

mean_score <- mean_score/100
```
```{r}
#Get best number of genes -> best mean p-value
N <- 3*(length(immune_metagenes) - 9)
DMFS_group <- mean_score[1:N]; p_value <- mean_score[N: (N+ N/3)]

min_index = which.min(p_value)
print(paste("The best N is: ", min_index + 9, ", with p-value of ", min(p_value)))
DMFS_group[((min_index*3) - 2):(min_index*3)]
```
Define the metagene
```{r}
ranked_genes <- univariate_Cox(exp_doubleneg, immune_metagenes)
metagene <- names(ranked_genes)[0:(min_index + 9)]

metagene
```
Test refined metagene
```{r}
all_data <- exp_doubleneg

#Find tertiles, assign low,medium,high to each group
all_data$sum <- rowMeans(all_data[,metagene])
tert = quantile(all_data$sum, c(0:3/3))
all_data$group = with(all_data, cut(sum, tert, include.lowest = T, labels = c("Low", "Medium", "High")))

#log-rank test
fit <- survfit(Surv(DMFSVal, DMFS)~ group, data = all_data)
print(autoplot(fit,  conf.int = F, ylim = c(0, 1), xlab = 'Time(years)', ylab = "DMFS(proportion)"))
surv_diff <- survdiff(Surv(DMFSVal, DMFS)~ group, data = all_data)
p_value <- 1 - pchisq(surv_diff$chisq, length(surv_diff$n)-1, lower.tail = TRUE)

print(p_value)
```

Refine metagene for HER+ patients
```{r}
immune_metagenes <- c("CXCL13","PLEK","IFNG","SLAMF7","IL2RB","PRF1","IRF1","PTPN22","IKZF1","APOBEC3G","IL2RA","ITGAL","CXCL9","GZMA","GZMB","HLA.E","CCR5","CD8A","SIRPG","CST7","GNLY","CECR1","PNOC","LCP1","HLA.DMB","GIMAP5","SEL1L3", "FGL2", "BIN2")

names(exp_HERpos)[names(exp_HERpos) == "HLA-E"] <- "HLA.E"
names(exp_HERpos)[names(exp_HERpos) == "HLA-DMB"] <- "HLA.DMB"
#Perform analysis 100 and take mean of DMFS percentage and p-values
mean_score <- c(0)
for (i in 1:100){
  mean_score <- mean_score + refine_metagene(exp_HERpos, immune_metagenes)
}

mean_score <- mean_score/100
```

```{r}
#Get best number of genes -> best mean p-value
N <- 3*(length(immune_metagenes) - 9)
DMFS_group <- mean_score[1:N]; p_value <- mean_score[N: (N+ N/3)]

min_index = which.min(p_value)
print(paste("The best N is: ", min_index + 9, ", with p-value of ", min(p_value)))
DMFS_group[((min_index*3) - 2):(min_index*3)]
```
Define the metagene
```{r}
ranked_genes <- univariate_Cox(exp_HERpos, immune_metagenes)
metagene <- names(ranked_genes)[0:(min_index + 9)]

metagene
```

Test refined metagene
```{r}
names(exp_HERpos)[names(exp_HERpos) == "HLA-E"] <- "HLA.E"
all_data <- exp_HERpos

#Find tertiles, assign low,medium,high to each group
all_data$sum <- rowMeans(all_data[,metagene])
tert = quantile(all_data$sum, c(0:3/3))
all_data$group = with(all_data, cut(sum, tert, include.lowest = T, labels = c("Low", "Medium", "High")))

#log-rank test
fit <- survfit(Surv(DMFSVal, DMFS)~ group, data = all_data)
print(autoplot(fit,  conf.int = F, ylim = c(0, 1), xlab = 'Time(years)', ylab = "DMFS(proportion)"))
surv_diff <- survdiff(Surv(DMFSVal, DMFS)~ group, data = all_data)
p_value <- 1 - pchisq(surv_diff$chisq, length(surv_diff$n)-1, lower.tail = TRUE)

print(p_value)
```

Define and test the consensus T-cell related metagene for double negative and HER+
```{r}
CTM <- c("CXCL13","PRF1", "IRF1", "IKZF1", "GZMB", "HLA.E")
all_data <- rbind(exp_doubleneg,exp_HERpos)

#Find tertiles, assign low,medium,high to each group
all_data$sum <- rowMeans(all_data[,CTM])
tert = quantile(all_data$sum, c(0:3/3))
all_data$group = with(all_data, cut(sum, tert, include.lowest = T, labels = c("Low", "Medium", "High")))

#log-rank test
DMFS <- data.frame(all_data$DMFS, row.names = rownames(all_data))
DMFS.Val <- data.frame(all_data$DMFSVal, row.names = rownames(all_data))
fit <- survfit(Surv(DMFSVal, DMFS)~ group, data = all_data)
print(autoplot(fit,  conf.int = F, ylim = c(0, 1), xlab = 'Time(years)', ylab = "DMFS(proportion)"))
surv_diff <- survdiff(Surv(DMFSVal, DMFS)~ group, data = all_data)
p_value <- 1 - pchisq(surv_diff$chisq, length(surv_diff$n)-1, lower.tail = TRUE)

print(p_value)
```
Only double_neg
```{r}
CTM <- c("CXCL13","PRF1", "IRF1", "IKZF1", "GZMB", "HLA.E")
all_data <- exp_doubleneg

#Find tertiles, assign low,medium,high to each group
all_data$sum <- rowMeans(all_data[,CTM])
tert = quantile(all_data$sum, c(0:3/3))
all_data$group = with(all_data, cut(sum, tert, include.lowest = T, labels = c("Low", "Medium", "High")))


#log-rank test
DMFS <- data.frame(all_data$DMFS, row.names = rownames(all_data))
DMFS.Val <- data.frame(all_data$DMFSVal, row.names = rownames(all_data))
fit <- survfit(Surv(DMFSVal, DMFS)~ group, data = all_data)
print(autoplot(fit,  conf.int = F, ylim = c(0, 1), xlab = 'Time(years)', ylab = "DMFS(proportion)"))
surv_diff <- survdiff(Surv(DMFSVal, DMFS)~ group, data = all_data)
p_value <- 1 - pchisq(surv_diff$chisq, length(surv_diff$n)-1, lower.tail = TRUE)

print(p_value)
```

Only HER+
```{r}
CTM <- c("CXCL13","PRF1", "IRF1", "IKZF1", "GZMB", "HLA.E")
names(exp_HERpos)[names(exp_HERpos) == "HLA-E"] <- "HLA.E"
all_data <- exp_HERpos

#Find tertiles, assign low,medium,high to each group
all_data$sum <- rowMeans(all_data[,CTM])
tert = quantile(all_data$sum, c(0:3/3))
all_data$group = with(all_data, cut(sum, tert, include.lowest = T, labels = c("Low", "Medium", "High")))


#log-rank test
DMFS <- data.frame(all_data$DMFS, row.names = rownames(all_data))
DMFS.Val <- data.frame(all_data$DMFSVal, row.names = rownames(all_data))
fit <- survfit(Surv(DMFSVal, DMFS)~ group, data = all_data)
print(autoplot(fit,  conf.int = F, ylim = c(0, 1), xlab = 'Time(years)', ylab = "DMFS(proportion)"))
surv_diff <- survdiff(Surv(DMFSVal, DMFS)~ group, data = all_data)
p_value <- 1 - pchisq(surv_diff$chisq, length(surv_diff$n)-1, lower.tail = TRUE)

print(p_value)
print(DMFS.groups)
```