---
title: "Metagene identification"
output: html_notebook
---
Add a new chunk by pressing 'Alt, Ctrl, i'

Load the generic dataset for metagene identification
```{r}
library("Biobase")
set.seed(123)
load(file = '../data/Generic.rda')

data <- generic_normalized
expression <- t(exprs(data))

#Randomly split the data into discovery and validation subset of equal size
split1<- sample(c(rep(0, 0.5 * nrow(expression)), rep(1, 0.5 * nrow(expression))))
train <- expression[split1== 0,]
test <- expression[split1 == 1,]

dim(train)
```

Cluster the genes using hierarchical clustering with linkage = complete (default) and distance measure =  pearson correlation on the discovery set (train)
```{r}
#Compute pearson correlation and apply hierarchical clustering
pear_cor <- as.dist(1-abs(cor(train, method = "pearson")))
dend <- hclust(pear_cor)

#Cut dendrogram at corr = 0.6 since dist
gene_clusters <- cutree(dend, h = 0.6)
```
```{r}
plot(dend)
```

```{r}
#Get cluster of genes (with at least 25 genes)
size_cluster <-table(unlist(gene_clusters))
cluster_interest <- rownames(size_cluster[size_cluster  >= 25])

gene_interest <- vector(mode='list', length=length(cluster_interest))
for (i in 1:length(cluster_interest)){
  genes <- names(gene_clusters[gene_clusters == cluster_interest[i]])
  gene_interest[[i]] <- genes
}
length(gene_interest) 
```
Re-Cluster using same method each cluster and only keep clusters that retained at least 20 genes from original set.
```{r}
final_gene_interest <- vector(mode='list')
u = 1 #current size of final_gene_interest
for (i in 1:length(gene_interest)){
  #Get data with only genes from current re-clustered set
  clust = test[,gene_interest[[i]]]
  
  #Compute pearson correlation and apply hierarchical clustering
  pear_cor <- as.dist(1-abs(cor(clust, method = "pearson")))
  dend <- hclust(pear_cor)
  
  #Cut dendrogram at corr = 0.4
  gene_clusters <- cutree(dend, h = 0.6)
  
  #Get cluster of genes (with at least 20 genes)
  size_cluster <- table(unlist(gene_clusters))
  cluster_interest <- names(size_cluster[size_cluster  >= 20])
  
  if (length(cluster_interest)!= 0){
  
    for (j in 1:length(cluster_interest)){
      genes <- names(gene_clusters[gene_clusters == cluster_interest[j]])
      final_gene_interest[[u]] <- genes
      u <- u + 1
    }
  }
}
final_gene_interest
```

Gene ontology analysis
```{r}
library(enrichR)
setEnrichrSite("Enrichr") # Human genes
websiteLive <- TRUE

dbs <- c("GO_Molecular_Function_2015", "GO_Cellular_Component_2015", "GO_Biological_Process_2015")
if (websiteLive) {
    enriched <- enrichr(final_gene_interest[[18]], dbs)
}
if (websiteLive) plotEnrich(enriched[[3]], showTerms = 20, numChar = 40, y = "Count", orderBy = "P.value")

```

